Legends for supplementary material

### 2019-05-14 Manuscript/Figures/figure_legends.Rmd
reordering and relabeling figures and tables

**Figure 5. Comparing ensemble system specific model performances for GM12878.** Performances of random forest classification models built on each of the 3 predictor types: OC (A), OP (B), and Distance (C). Within each predictor type, and across resolution, MCCs metrics were recorded for models subsequently implemented on re-sampled data using the 4 re-sampling techniques: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Figure 6. Comparing annotation-specific random forest models implemented on SMOTE re-sampled data with distance-type predictors for GM12878.** MCC performances were compared for models using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Figure 8. Predictive importance** A heatmap illustrating the importance of the top 20 most predictive transcription factors for random forest models implemented on SMOTE resampled data with distance-type predictors for GM12878. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

@@

<!--
**Figure 4. Barplots comparing model performance for each predictor type.** Model performances were evaluated using F1-score, MCC, and balanced accuracy for each of the 4 resampling techniques, across both the GM12878 (A) and K562 (B) cell lines. The results presented are for models applied to 10 kb resolution data.

**Figure 5. Line plots comparing models using SMOTE resampling with distance-type predictors.** Model performances were evaluated using F1-score, MCC, and balanced accuracy for each of the 4 data resolutions, across both the GM12878 (red, solid line) and K562 (blue; dashed line) cell lines. For a given metric, each point represents a model applied to a particular resolution with distance type predictors for the feature space, after which the SMOTE algorithm was used to create balanced classes on the trianing set.

**Figure 6. Variable importance of the top 20 most predictive genomic annotations for models using SMOTE resampling with distance-type predictors.** The list of genomic features are the union of the top 20 most important genomic annotations across resolutions for the GM12878 (A) and K562 (B) cell lines. Variable importance is assigned on a continuous scale (range: 0-100). Grid entries filled in gray denote an annotation that was not featured as one of the top 20 most predictive variables in a given model. The genomic annotations are sorted by highest level of importance obtained from a random forest model using distance-type predictors at the 10 kb data resolution, while using SMOTE resampling, for each cell line respectively.
-->

<!--
**Figures 4-8. Precision-Recall curves for 10 kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on 10 kb resolution data, across each resampling technique. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). 
-->

<!--
**Figure 9-13. Recursive Feature Elimination for 10 kb data** Beginning with the full set of distance type predictors for each cell line, models built on 10 kb resolution data were recursively evaluated at different subsets of feature size. For convenience and computational efficiency, feature size decreasing in powers of 2 was considered. The maximum size for feature subsets differed between cell line due the available annotations that were considered and provided in Supplementary Table 1. Near optimal performance was achieved using the top 32 distance type predictors for both the GM12878 (red) and K562 cell lines (blue).
-->

>>>>

**Figure 4. Comparing ensemble system specific model performances for GM12878.** Performances of random forest classification models built on each of the 3 predictor types: OC (A), OP (B), and Distance (C). Within each predictor type, and across resolution, MCCs metrics were recorded for models subsequently implemented on re-sampled data using the 4 re-sampling techniques: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Figure 5. Comparing annotation-specific random forest models implemented on SMOTE re-sampled data with distance-type predictors for GM12878.** MCC performances were compared for models using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Figure 6. Predictive importance** A heatmap illustrating the importance of the top 20 most predictive transcription factors for random forest models implemented on SMOTE resampled data with distance-type predictors for GM12878. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

@@

### 2019-05-14 Manuscript/Supplementary/supplementary_legends.Rmd
reordering and relabeling figures and tables

**Supplementary Table 4. Hi-C data.** A complete list of the Hi-C data libraries used to obtain the locations of topologically associating domains.

**Supplementary Table 8. Summary statistics for called TADs.** The following summary statistics were recorded regarding the average number and length of TADs (along with standard deviations) that were called using the Arrowhead algorithm [@durand2016juicebox], across each of the autosomal chromosomes (excluding chromosome 9).

@@

**Supplementary Figure 10. Comparing ensemble system specific model performances for K562.** Performances of random forest classification models built on each of the 3 predictor types: OC (A), OP (B), and Distance (C). Within each predictor type, and across resolution, MCCs metrics were recorded for models subsequently implemented on re-sampled data using the 4 re-sampling techniques: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Supplementary Table 4. Ranking predictive importance between predictor types for annotation-specific models for GM12878.** Annotation-specific random forest classification models were implemented on SMOTE resampled data, across resolution. Each model contained each of the 3 predictor types for one genomic annotation. Predictive importance was rescaled to values between 0 and 100. The predictor type with importance value of 100 indicates it was the most informative predictor type out of the 3.

**Supplementary Table 5. Ranking predictive importance between predictor types for annotation-specific models for K562.** Annotation-specific random forest classification models were implemented on SMOTE resampled data, across resolution. Each model contained each of the 3 predictor types for one genomic annotation. Predictive importance was rescaled to values between 0 and 100. The predictor type with importance value of 100 indicates it was the most informative predictor type out of the 3.

**Supplementary Figure 11. Comparing annotation-specific random forest models implemented on SMOTE re-sampled data with distance-type predictors for K562.** MCC performances were compared for models using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Supplementary Figure 15. Predictive importance** A heatmap illustrating the importance of the top 20 most predictive transcription factors for random forest models implemented on SMOTE resampled data with distance-type predictors for K562. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

**Supplementary Figure 12. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 25 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 13. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 50 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 14. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 100 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 15. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 10 kb resolution, on K562.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 16. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 25 kb resolution, on K562.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 17. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 50 kb resolution, on K562.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

**Supplementary Figure 18. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 100 kb resolution, on K562.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.

>>>>

**Supplementary Table 1. Hi-C data.** A complete list of the Hi-C data libraries used to obtain the locations of topologically associating domains.

@@

**Supplementary Figure 3. Comparing ensemble system specific model performances for K562.** Performances of random forest classification models built on each of the 3 predictor types: OC (A), OP (B), and Distance (C). Within each predictor type, and across resolution, MCCs metrics were recorded for models subsequently implemented on re-sampled data using the 4 re-sampling techniques: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Supplementary Figure 4. Comparing annotation-specific random forest models implemented on SMOTE re-sampled data with distance-type predictors for K562.** MCC performances were compared for models using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Supplementary Figure 5. Predictive importance** A heatmap illustrating the importance of the top 20 most predictive transcription factors for random forest models implemented on SMOTE resampled data with distance-type predictors for K562. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

<!--
**Supplementary Figure 12. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 25 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.
-->
